
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% This is the template for submission to ISCA 2016
% The cls file is a modified from  'sig-alternate.cls'
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass{sig-alternate}
%\usepackage{mathptmx} % This is Times font

%\newcommand{\ignore}[1]{}
\usepackage{fancyhdr}
\usepackage[normalem]{ulem}
\usepackage[hyphens]{url}
\usepackage{hyperref}

\usepackage{tabularx}
\usepackage{ulem}
%\usepackage{program}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% I need these so it compiles on my Ubuntu machine
% There's some issue with the algorithm2e package
\makeatletter
\newif\if@restonecol
\makeatother
\let\algorithm\relax
\let\endalgorithm\relax
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage[boxruled]{algorithm2e}


%\usepackage{algorithm2e}
%\usepackage{mathptmx} % This is Times font

%%%%%%%%%%%---SETME-----%%%%%%%%%%%%%
\newcommand{\microsubmissionnumber}{266}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\fancypagestyle{firstpage}{
  \fancyhf{}
\setlength{\headheight}{50pt}
\renewcommand{\headrulewidth}{0pt}
  \fancyhead[C]{\normalsize{MICRO 2016 Submission
      \textbf{\#\microsubmissionnumber} -- Confidential Draft -- DO NOT DISTRIBUTE}}
  \pagenumbering{arabic}
}

%%commands
\input{macros}
\newcommand{\TODO}[1]{\textcolor{red}{\todo{#1}}}
\renewcommand{\arraystretch}{1.5}

%%%%%%%%%%%---SETME-----%%%%%%%%%%%%%
\title {MemPod: A Clustered Architecture for Efficient and Scalable Migration in Flat Address Space Multi-level Memories}
\author{}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}
\maketitle
\thispagestyle{firstpage}
\pagestyle{plain}

\begin{abstract}

In the near future die-stacked DRAM will be present alongside off-chip memories in hybrid memory systems. A large body of recent research has explored possible uses and performance-improving mechanisms regarding such a memory configuration. The first approach is to use this new memory as a last level cache and it has been shown to be beneficial for system performance. Another approach is to use the on-chip memory as an extension of the off-chip one in a ``flat address space'' configuration, exposing more memory capacity to applications leading to performance improvement under capacity constrained workloads.

In this paper we propose MemPod, a scalable and efficient memory management mechanism for flat address space hybrid memories. MemPod monitors main memory activity and upon predefined intervals migrates the most frequently accessed memory pages to the faster on-chip memory. We propose a novel architectural organization and the use of ``Majority Element Algorithm'' for activity tracking. 

Our results with multi-programmed workloads show MemPod to improve Average Main Memory Time (AMMT) by up to 29\% and 11\% on average compared to the state-of-the-art, while our analysis shows that MemPod is the most viable option compared to the state of the art, as memories become faster in the future. MemPod's novel activity tracking approach leads to significant cost reduction ($\sim$6000x lower space requirements) and incredible improvement on future prediction accuracy (58\% average improvement) over traditionally used full counters. 

\end{abstract}

\noindent \textbf{Keywords:} Memory architecture, Die-stacked memory

\input{sections/Introduction.tex}
\input{sections/Background.tex}
%\input{sections/RelatedWork.tex}
\input{sections/MEA.tex}
\input{sections/Architecture.tex}
\input{sections/Results.tex}
\input{sections/Conclusions.tex}

%\bstctlcite{bstctl:etal, bstctl:nodash, bstctl:simpurl}
%\bibliographystyle{IEEEtranS}
\bibliographystyle{ieeetr}
\bibliography{references}

\end{document}
